How to compute the fusion product of MV cycles in type A

### 2021-04-29 draft.tex
meeting notes

The variety $X(\tau)$ defined by
% from $\tau\in YT(\lambda)_\mu$ by 
\[
X(\tau) = \overline{
    \{
        A \in  \TT_\mu\cap\n : A\big|_{\CC^{|\mu(i)|}} \in \OO^{\lambda(i)} \text{ for each } i = 1,\dots,m
    \}
}    
\]
is called a \textbf{generalized orbital variety}. 
% 
It inherits a crystal datum called a \textbf{Lusztig datum} from its tableau. 
\acom{Cite Joel? Hide dependence on reduced expression?}
The Lusztig datum $n_\bullet(\tau)$ of the tableau $\tau$ is got from its \textbf{Gelfand--Tsetlin pattern} $\gt(\tau) = (\lambda(i)_j)_{1\le j\le i\le m}$ \acom{cite BZ} by the formula 
$$
n_\bullet(\tau)_{i,j} = \lambda(i+1)_j - \lambda(i)_j
$$
for $(i,j) = (1,1), (2,1), \dots, (m,1), (2,2),\dots,(m,2),\dots,(m,m)$. The pattern is recorded as a lower-triangular matrix (array of shapes of subtableaux) and the datum is recorded as a sequence.
% 
For example, if 
\[
    \tau = \young(1112,23) 
\]
then
\begin{gather*}
%    \text{if } \tau = \young(1112,23) \text{ then } 
\gt(\tau) = \begin{matrix}
       3 \\
       4 & 1 \\ 
       4 & 2 & 0 
   \end{matrix} \text{ and }
   n_\bullet(\tau) = (1,0,1)\,. 
\end{gather*}
% 
% TODO
% \begin{theorem}(CITE KAM07)
%     Taking $\vi$-Luszig datum is a bijection $P\to\NN^{\ell}$. 
% \end{theorem}
% 
We need the following two results from \cite{dthesis}.
\begin{proposition}
    \label{pr:govsasirrecs}
    The decomposition
    \begin{equation*}
        % \label{eq:govsasirrecs}
        \overline{\OO}^\lambda\cap\TT_\mu\cap\n = \bigsqcup_{\tau\in YT(\lambda)_\mu} X(\tau)
    \end{equation*}
    is\dots
\end{proposition}
% 
Denote by $Z(\tau)\subset\Gr$ the closure of the image of $X(\tau)$ under the isomorphism of \Cref{cor:mvy}. 
\acom{Still need Lusztig data of MV cycles; with it $B(\infty)$, stable vs ordinary, etc?}
\jcom{Do you want $ Z(\tau)$ to be the closure?}
\acom{Sure}

% 
\begin{proposition}
    \label{prop:dth}
    % The isomorphism 
    % $\phi: \overline{\OO}^\lambda\cap\TT_\mu\cap\n \to \overline\Gr^\lambda\cap S^\mu_-$ 
    % is compatible with the decomposition of \Cref{pr:govsasirrecs}
    % the bijection $YT(\lambda)_\mu \to \irr \overline{\OO}^\lambda\cap\TT_\mu\cap\n$ 
    % in the sense that 
    $Z(\tau)$ is dense in the MV cycle whose Lusztig datum is $n_\bullet(\tau)$. 
\end{proposition}
\acom{We would like to use the action of $W$ which gives isomorphisms
$Gr^\lambda \cap Gr_\mu \cap S^\mu \cong Gr^\lambda \cap Gr_{w\mu} \cap S^{w\mu}$ to generalize to non-dominant $\mu$. But the Lusztig datum of $\tau\in YT(\lambda)_\mu$ may not appear as the Lusztig datum of any element of $YT(\lambda)_{w\mu}$..}
% This gives us the dimension of Gr^\lambda \cap Gr_\mu \cap S^\mu.
\begin{lemma}
Alternate description. Works for non-dominant $\mu$. 
$$
Z(\tau) = \{
L\in\Gr : L\cap\CC[t]^i\in\Gr^{\lambda(i)}\cap S^{\mu(i)}\text{ for all }i \}
$$
is an MV cycle. 
\end{lemma}
% Given $(\tau',\tau'')\in YT(\lambda')_{\mu'}\times YT(\lambda'')_{\mu''}$ denote by $\lambda'(i),\lambda''(i)$ (resp.\ $\mu'(i),\mu''(i)$) the shape (resp.\ the weight) of $\tau'(i),\tau''(i)$ where $\tau^?(i)$ denotes the tableau got from $\tau^?$ by discarding all 
% % but its first $i$ boxes
% boxes of weight exceeding $i$. 
% % If the tableau contains repeated entries then first means leftmost.
% 
\jcom{I don't think that you should mention this partial inverse here.  I think that you should explain here how the generalized orbital varieties are indexed by the tableaux, in other words, how $ X(\tau) $ is defined.}
\acom{Ok, I moved it.}

@@

\begin{proof}
    % We wish to apply \Cref{prop:dth}. Nope
    % We will also need \Cref{le:Grl1l2}. Nope 
    % Consider the following diagram.
    % \[
    % \begin{tikzcd}
    %     X'\ast_s X'' \ar[r,dashed,"?"] \ar[d,"\bar g"] & X'\times X'' \ar[d," \phi\times \phi"] \\ 
    %     \Gr_{0,s}\ar[r,"\tau_s"] & \Gr\times\Gr
    % \end{tikzcd}
    % \]
    % \acom{Not sure what's a good notation for the vertical arrows. Don't want to introduce $\phi$. $\bar g(A) = g(A)\CC[t]^m$ temporarily.}
    % We show that the image of $\tau_s\circ \bar g$ coincides with the image of $\phi\times\phi$ which is dense in $Z'\times Z''$ by \Cref{prop:dth}.
    % % 
    Let $A\in X(\tau')\ast_s X(\tau'')$, let $i = 1,\dots, m$.  
    % Denote by 
    % $E_a$ the generalized $a$-eigenspace of $A$, and by 
    Clearly, the lattices $L=g(A)\CC[t]^m$ and $l=g(A\big|_{\CC^{|\mu(i)|}})\CC[t]^i$ are related by
    $$
    L\cap \CC[t]^i = l. 
    $$
    By definition of $X'\ast_s X''$,  $A\big|_{\CC^{|\mu(i)|}}$ has Jordan type $((\lambda'(i),0),(\lambda''(i),s))$ and lies in $\UU^{\mu'(i),\mu''(i)}_{0,s}$. So, by \Cref{th:OTGrW}, $l\in\Gr_{0,s}^{\lambda'(i),\lambda''(i)}\cap S^{\mu'(i),\mu''(i)}_{0,s}$ which fits the alternate description of $Z(\tau)$..

>>>>

The variety $X(\tau)$ defined by taking the top dimensional component of \acom{todo}
% from $\tau\in YT(\lambda)_\mu$ by 
\[
% X(\tau) = 
\overline{
    \{
        A \in  \TT_\mu\cap\n : A\big|_{\CC^{|\mu(i)|}} \in \OO^{\lambda(i)} \text{ for each } i = 1,\dots,m
    \}
}    
\]
is called a \textbf{generalized orbital variety}. 
% 
It inherits a crystal datum called a \textbf{Lusztig datum} from its tableau. 
\acom{Cite Joel? Hide dependence on reduced expression?}
The Lusztig datum $n_\bullet(\tau)$ of the tableau $\tau$ is got from its \textbf{Gelfand--Tsetlin pattern} $\gt(\tau) = (\lambda(i)_j)_{1\le j\le i\le m}$ \acom{cite BZ} by the formula 
$$
n_\bullet(\tau)_{i,j} = \lambda(i+1)_j - \lambda(i)_j
$$
for $(i,j) = (1,1), (2,1), \dots, (m,1), (2,2),\dots,(m,2),\dots,(m,m)$. The pattern is recorded as a lower-triangular matrix (array of shapes of subtableaux) and the datum is recorded as a sequence.
% 
For example, if 
\[
    \tau = \young(1112,23) 
\]
then
\begin{gather*}
%    \text{if } \tau = \young(1112,23) \text{ then } 
\gt(\tau) = \begin{matrix}
       3 \\
       4 & 1 \\ 
       4 & 2 & 0 
   \end{matrix} \text{ and }
   n_\bullet(\tau) = (1,0,1)\,. 
\end{gather*}
% 
% TODO
% \begin{theorem}(CITE KAM07)
%     Taking $\vi$-Luszig datum is a bijection $P\to\NN^{\ell}$. 
% \end{theorem}
% 
We need the following two results from \cite{dthesis}.
\begin{proposition}
    \label{pr:govsasirrecs}
    The set of.. is the irrecs of.. 
    \begin{equation*}
        % \label{eq:govsasirrecs}
        \overline{\OO}^\lambda\cap\TT_\mu\cap\n = \bigsqcup_{\tau\in YT(\lambda)_\mu} X(\tau)
    \end{equation*}
    is\dots
\end{proposition}
% 
Denote by $Z(\tau)\subset\Gr$ the closure of the image of $X(\tau)$ under the isomorphism of \Cref{cor:mvy}. 
\acom{Still need Lusztig data of MV cycles; with it $B(\infty)$, stable vs ordinary, etc?}
\jcom{Do you want $ Z(\tau)$ to be the closure?}
\acom{Sure}

% 
\begin{proposition}
    \label{prop:dth}
    % The isomorphism 
    % $\phi: \overline{\OO}^\lambda\cap\TT_\mu\cap\n \to \overline\Gr^\lambda\cap S^\mu_-$ 
    % is compatible with the decomposition of \Cref{pr:govsasirrecs}
    % the bijection $YT(\lambda)_\mu \to \irr \overline{\OO}^\lambda\cap\TT_\mu\cap\n$ 
    % in the sense that 
    $Z(\tau)$ is dense in the MV cycle whose Lusztig datum is $n_\bullet(\tau)$. 
\end{proposition}
\acom{We would like to use the action of $W$ which gives isomorphisms
$Gr^\lambda \cap Gr_\mu \cap S^\mu \cong Gr^\lambda \cap Gr_{w\mu} \cap S^{w\mu}$ to generalize to non-dominant $\mu$. But the Lusztig datum of $\tau\in YT(\lambda)_\mu$ may not appear as the Lusztig datum of any element of $YT(\lambda)_{w\mu}$..}
% This gives us the dimension of Gr^\lambda \cap Gr_\mu \cap S^\mu.
\begin{lemma}
Alternate description. Works for non-dominant $\mu$. 
$$
Z(\tau) = \overline{\{
L\in S^\mu : L\cap\CC[t]^i\in\Gr^{\lambda(i)}\text{ for all }i \}}
$$
is an MV cycle. 
\end{lemma}
% Given $(\tau',\tau'')\in YT(\lambda')_{\mu'}\times YT(\lambda'')_{\mu''}$ denote by $\lambda'(i),\lambda''(i)$ (resp.\ $\mu'(i),\mu''(i)$) the shape (resp.\ the weight) of $\tau'(i),\tau''(i)$ where $\tau^?(i)$ denotes the tableau got from $\tau^?$ by discarding all 
% % but its first $i$ boxes
% boxes of weight exceeding $i$. 
% % If the tableau contains repeated entries then first means leftmost.
% 
\jcom{I don't think that you should mention this partial inverse here.  I think that you should explain here how the generalized orbital varieties are indexed by the tableaux, in other words, how $ X(\tau) $ is defined.}
\acom{Ok, I moved it.}

@@

\begin{proof}
    % We wish to apply \Cref{prop:dth}. Nope
    % We will also need \Cref{le:Grl1l2}. Nope 
    % Consider the following diagram.
    % \[
    % \begin{tikzcd}
    %     X'\ast_s X'' \ar[r,dashed,"?"] \ar[d,"\bar g"] & X'\times X'' \ar[d," \phi\times \phi"] \\ 
    %     \Gr_{0,s}\ar[r,"\tau_s"] & \Gr\times\Gr
    % \end{tikzcd}
    % \]
    % \acom{Not sure what's a good notation for the vertical arrows. Don't want to introduce $\phi$. $\bar g(A) = g(A)\CC[t]^m$ temporarily.}
    % We show that the image of $\tau_s\circ \bar g$ coincides with the image of $\phi\times\phi$ which is dense in $Z'\times Z''$ by \Cref{prop:dth}.
    % % 
    Let $A\in X(\tau')\ast_s X(\tau'')$, let $i = 1,\dots, m$.  
    % Denote by 
    % $E_a$ the generalized $a$-eigenspace of $A$, and by 
    Clearly, the lattices $L=g(A)\CC[t]^m$ and $l=g(A\big|_{\CC^{|\mu(i)|}})\CC[t]^i$ are related by
    $$
    L\cap \CC[t]^i = l. 
    $$
    By definition of $X'\ast_s X''$,  $A\big|_{\CC^{|\mu(i)|}}$ has Jordan type $((\lambda'(i),0),(\lambda''(i),s))$ and lies in $\UU^{\mu'(i),\mu''(i)}_{0,s}$. So, by \Cref{th:OTGrW}, $l\in\Gr_{0,s}^{\lambda'(i),\lambda''(i)}$ which fits the alternate description of $Z(\tau)$.
